Automatic Matching of Bullet Lands

### 2016-01-04 imaging-paper.Rnw
little blurb on smoothing factor

\subsection{Varying smoothing factor}\label{sec:smoothing}
As mentioned earlier, the algorithm for detecting peaks and valleys depend on the selection of a smoothing window, called the smoothing factor. For detecting and removing the grooves prior to fitting a loess regression we selected a smoothing factor of 35, while for detecting the peaks/valleys of the loess residuals a smoothing factor of 25 seems more appropriate. %We can investigate the performance of peak/valley detection across different smoothing factors.
\hh{XXX the smoothing factor needs a few more details: a smoothing factor of 5 means that we consider the two measurements on the left and the right of the observations under consideration XXX}
Figure~\ref{fig:varysmooth} displays the  peaks and valleys detected in the same signature at smoothing factors of 5, 25, and 45, respectively. The dark line corresponds to the smoothed values, while the grey line in the back shows the raw signature. The choice of smoothing factor is a classical decision of a bias/variance trade-off. It is immediately clear that a small smoothing factor like 5 is a poor choice. It results in a significant amount of noise in the data such that even just a point or two can skew the rolling average enough for a peak or valley to be detected. Given that the striation patterns are typically much larger, we are in effect muddying the waters by performing such minimal smoothing. Another consideration is that the smoothing should not fall below the  resolution of the equipment at which the surface measurements are taken -- in order to not introduce artefacts in the analysis.

>>>>

\subsection{Varying smoothing factor}\label{sec:smoothing}
As mentioned earlier, the algorithm for detecting peaks and valleys depend on the selection of a smoothing window, called the smoothing factor or span. A smoothing factor of $k$ means that the  $k$ closest observations to $x_o$ are considered for a fit for $x_o$. Because surface measurements are recorded at an equidistand resolution (here, of 1.5625$\mu m$), we decided to only consider odd smoothing factors $2k + 1$, which means that the  $k$ observations to the left and right of $x_o$ are considered for a local fit of $x_o$..  For detecting and removing the grooves prior to fitting a loess regression we selected a smoothing factor  of 35, while for detecting the peaks/valleys of the loess residuals a smoothing factor of 25 seems more appropriate. %We can investigate the performance of peak/valley detection across different smoothing factors.

Figure~\ref{fig:varysmooth} displays the  peaks and valleys detected in the same signature at smoothing factors of 5, 25, and 45, respectively. The dark line corresponds to the smoothed values, while the grey line in the back shows the raw signature. The choice of smoothing factor is a classical decision of a bias/variance trade-off. It is immediately clear that a small smoothing factor like 5 is a poor choice. It results in a significant amount of noise in the data such that even just a point or two can skew the rolling average enough for a peak or valley to be detected. Given that the striation patterns are typically much larger, we are in effect muddying the waters by performing such minimal smoothing. Another consideration is that the smoothing should not fall below the  resolution of the equipment at which the surface measurements are taken -- in order to not introduce artefacts in the analysis.
